Natural Language Interfaces to Structured Data

### 2012-07-28 trunk/tr-nlisurvey/src/nli-survey.tex
section 2

% added packages
\usepackage{multirow}
\usepackage{url}

@@

\ifCLASSINFOpdf
\usepackage[pdftex]{graphicx}
% declare the path(s) where your graphic files are
\graphicspath{{images/}}
% and their extensions so you won't have to specify these with
% every instance of \includegraphics
\DeclareGraphicsExtensions{.png}
\else

@@

\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand{\TODO}[1]{{\textcolor{green}{TODO:#1}}}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Natural Language Interfaces to Structured Data}

@@

\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Abstract will be written later\ldots
\end{abstract}

\begin{keywords}
Computer Society, IEEEtran, journal, \LaTeX, paper, template.
\end{keywords}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  INTRODUCTION                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}

%{\bf Possible quotation:} M.P. Schutzenberger, 1970: (in the AI domain: humans
%learn how to use systems, see the usage of Google, keywords etc.)

Natural Language (NL) is the most natural and easy way to express information
need~\cite{Hearst:2011:NSU:2018396.2018414}. Moreover, it is not restricted to
experts who have a good knowledge on how to formulate a formal query.

NL as an interface has been investigated by researchers in both Information
Retrieval (IR) and Database communities. In the IR community however, the
keyword paradigm has become the most popular, for sure because of the success of
commercial search engines. NL interfaces for unstructured documents are
refered to as Question Answering (Q\&A) systems, that have received much
researchers' attention as the Web had made avaialble huge amount of textual
data, and recently as the famous {\it Jeopardy!} quiz was won by IBM's {\sc
Watson}~\cite{FerrucciBCFGKLMNPSW10} system.
{\sc WolframAlpha}\footnote{\url{http://www.wolframalpha.com/}} is a popular
search service based on structured data, that accepts keyword queries as well
as some questions in natural language.

NL interfaces (to structured data) has focused much researchers' attention for
decades, but the field seems to have known a renewed interest for a few
years, probably thanks to the development of the semantic Web. {\bf See also
the quotation in~\cite{DBLP:journals/semweb/LopezUSM11}: users prefer NL interfaces
than logic used in the Semantic Web.}
However, the challenge of answering NL
questions is far from being solved~\cite{Kotov:2010:TNQ:1772690.1772746}.

@@

In this survey, we try to point out the major dimensions of existing systems
and to present the new trends and challenges of state-of-the-art systems.
The history of NL interfaces can be developped as follows:
\begin{enumerate}
  \item early years of domain-specific systems
  \item complex question answering in a specific domain
  \item rise of domain awareness in NL interfaces (through learning techniques)
  \item data-driven systems with respect to
  \begin{itemize}
    \item resources: almost domain-independent systems
    \item large-scale data: domain-independent systems
  \end{itemize}
\end{enumerate}

We will present the ``big picture'' of the approaches in
section~\ref{sec:big-picture}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BIG PICTURE                         %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Main Dimensions}
In this section, we introduce the main dimensions based on which NLI approaches can be distinguished. The input to every NLI is the (1) \emph{data}, 
and (2) \emph{user questions} representing information needs, which are translated to an (3) \emph{internal representation} of the needs employed by the system and/or directly mapped to 
(4) \emph{database queries} that finally, are executed by the underlying query engine to produce the final answers. The main problem tackled by a NLI approach is to transform the question to the internal representation and then to the database query (2 $\mapsto$ 3 $\mapsto$ 4), or to directly mapped the question to the query (2 $\mapsto$ 4). This problem is hard when considering the large and rapidly evolving mass of structured data that have become available. The major challenges modern NLI systems are facing in this regard is to operate across domains (5, \emph{domain-independence}) as well as to adopt to new domains (6, \emph{portability}). For evaluating the quality of the NLI output, different (6) \emph{metrics} based on the user questions the system can understand as well as the database queries it can produce, have been proposed.

----------
\TODO{old content that might be considered:}
Dimensions:
\begin{itemize}
  \item What kind of query? (NL, controlled language)
  \item What are the data? 
  \item Target structured query: select query? update query?
  \item What are the answers?
  \item How portable is the system? Does it improve its performance over time?
  \item What is the linguistic coverage of the system? 
  \item Error feedback: how well is the user informed of misunderstanding of her
  question
  \item Predictability: how the system disambiguates / lets the user decide what
  should be the correct interpretation
  \item Performance: response time / evaluation metrics
\end{itemize}

----------

\subsection{Data}
\label{sec:data}
\TODO{Discuss the different types of structured data models that have been used, make clear that while they vary in expressiveness, they commonly enables capturing resources/entities/objects and relationships between then. Besides domain information captured in terms of entities and relationships, also point out other popular types of information that have considered, e.g. temporal and spatial information}

Structured data:
\begin{itemize}
  \item relational databases; data warehouses
  \item XML databases
  \item early data structures (Prolog databases; list-structured databases)
  \item ontologies; {\it linked data}
\end{itemize}
%Data structured in a database belong to a {\it domain}\ldots
In the survey, we refer to {\it database elements} which stand for named
properties of the database. For instance in a relational database, they stand
for relation names, attributes names and values, \ldots

Data structures are associated with a data model or schema, which is explicitely
defined in modern databases.
Despite the effort of modeling and conceptualizing structured data, there is
still a gap between the physical view of structured data (database
administrators' view point) and the conceptual view of end users; interfaces
concerned in this survey attempt to bridge this gap.

\subsection{Question}
\TODO{Discuss different types user questions}

\label{sec:big-picture-question}
Traditional database management systems provide an interface that allows users
to query the data in a formal language (e.g., SQL).
\subsubsection{Keyword query}
A {\it keyword query} consists in a set of words
traditionally used in the context of document search (i.e. searching relevant
documents given a textual query).
The classic idea is to represent documents in a vector model, where the vector
is composed of terms of documents. Search engines have made this paradigm
popular.
\cite{Hearst:2011:NSU:2018396.2018414} reports that the number of words
in queries over Web search engines trends to increase (the experiment compared
queries performed over a month in 2009 and in 2010): queries from 5 to 8 words
increase up to 10\% while queries from 1 to 4 words decrease up to 2\%.
It seems indeed that users are more and more aware of new capabilities of
state-of-the-art search engines, and are less reluctant to express their
information need in a more ``natural'' way.

@@

\subsection{Internal Representation}
\label{sec:big-picture-syntactic}
\TODO{Discuss different types user questions}

@@

\subsection{Database Query}
\label{sec:big-picture-semantic}
\TODO{Discuss query languages; discuss the different types of queries expressed by these languages, e.g. entity queries, relational queries, additional constructs such as group by etc.}

@@

TOBECONSIDERED
-------
\subsubsection{Expressivity of database query}
\label{sec:big-picture-expressivity}
-----------

\subsection{Results}
\label{sec:big-picture-results}
NL interfaces produce database queries.
Those queries must be then be executed by underlying DBMS.
The execution of queries can lead to different failing states:
\begin{itemize}
  \item the query execution fails (the generated database query is not valid)
  \item the query execution lead to an empty result set
\end{itemize}
In these cases, the system may inform the user and/or suggest/rephrase the
query. This is especially performed in systems belonging to the feedback
approach (see section~\ref{sec:feedback-driven}).

\subsection{Query-independence}
\subsection{Portability}

>>>>

\usepackage{multirow}
\usepackage{url}
\usepackage{color}

@@

\ifCLASSINFOpdf
\usepackage[pdftex]{graphicx}
\graphicspath{{images/}}
\DeclareGraphicsExtensions{.png}
\else

@@

%\hyphenation{op-tical net-works semi-conduc-tor}

\newcommand\TODO[1]{{\textcolor{red}{TODO:#1}}}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Natural Language Interfaces to Structured Data}

@@

\IEEEcompsoctitleabstractindextext{
\begin{abstract}
Natural language interfaces to databases allow users to formulate queries in
natural language to databases. The benefits are twofold: first, users do not
have to know how the data have been modeled in the database, what term refers to
what database entity. Second, these systems can interface several data sources
that are possibly of different implementation and which data belong to different
domains.
In this paper we provide a survey of major natural language interfaces, from the
very beginning of artificial intelligence to most recent systems.
\end{abstract}

\begin{keywords}
Natural Language, Database
\end{keywords}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  INTRODUCTION                       %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}
Natural Language (NL) is the most natural and easy way to express information
need~\cite{Hearst:2011:NSU:2018396.2018414}. Moreover, it is not restricted to
experts who have a good knowledge on how to formulate a formal query.

NL as an interface has been investigated by researchers in both Information
Retrieval (IR) and Database communities. In the IR community however, the
keyword paradigm has become the most popular, for sure because of the success
of commercial search engines. 
%NL interfaces for unstructured documents are refered to as Question Answering
%(Q\&A) systems, that have received much researchers' attention as the Web had
Interfaces to unstructured documents has become popular as
the Web has made available huge amount of textual data, and recently as the
famous {\it Jeopardy!} quiz was won by IBM's {\sc
Watson}~\cite{FerrucciBCFGKLMNPSW10} system.
This trend seems to evolve slightly. For instance,
{\sc WolframAlpha}\footnote{\url{http://www.wolframalpha.com/}} is a popular
search service based on structured data, that accepts keyword queries as well
as some questions in natural language.

NL interfaces (to structured data) has focused much researchers' attention for
decades, but the field seems to have known a renewed interest for a few
years, probably thanks to the development of the semantic Web. 
Today's semantic technologies can be directly manipulated by experts, and
therefore there is still a need for interfaces. Indeed, users prefer NL
interfaces than the logic which is used in the semantic
Web~\cite{Kotov:2010:TNQ:1772690.1772746}.
However, the challenge of answering NL questions is far from being
solved~\cite{Kotov:2010:TNQ:1772690.1772746}.

@@

In this survey, we outline the major dimensions of existing systems
and present the new trends and challenges of state-of-the-art systems.
The history of NL interfaces can be developped as follows:
\begin{enumerate}
  \item early years of domain-specific systems
  \item complex question answering in a specific domain
  \item rise of domain awareness in NL interfaces (through learning techniques)
  \item data-driven systems with respect to
  \begin{itemize}
    \item resources: almost domain-independent systems
    \item large-scale data: domain-independent systems
  \end{itemize}
\end{enumerate}
We present section~\ref{sec:main-dimensions} the main dimensions used throughout
the paper.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BIG PICTURE                         %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Main Dimensions}
\label{sec:main-dimensions}
In this section, we introduce the main dimensions based on which NLI approaches
can be distinguished. The input to every NLI is the (1)~\emph{data}, and
(2)~\emph{user questions} representing information needs, which are translated
to an (3)~\emph{internal representation} of the needs employed by the system
and/or directly mapped to (4)~\emph{database queries} that finally, are
executed by the underlying query engine to produce the final answers. The main
problem tackled by a NLI approach is to transform the question to the internal
representation and then to the database query (2 $\mapsto$ 3 $\mapsto$ 4), or
to directly mapped the question to the query (2 $\mapsto$ 4). This problem is
hard when considering the large and rapidly evolving mass of structured data
that have become available. The major challenges modern NLI systems are facing
in this regard is to operate across domains (5)~\emph{domain-independence} as
well as to adopt to new domains (6)~\emph{portability}. For evaluating the
quality of the NLI output, different (7)~\emph{metrics} based on the user
questions the system can understand as well as the database queries it can
produce, have been proposed.

\TODO{old content that might be considered:} 
Dimensions:
\begin{itemize}
  \item What kind of query? (NL, controlled language)
  \item What are the data? 
  \item Target structured query: select query? update query?
  \item What are the answers?
  \item How portable is the system? Does it improve its performance over time?
  \item What is the linguistic coverage of the system? 
  \item Error feedback: how well is the user informed of misunderstanding of her
  question
  \item Predictability: how the system disambiguates / lets the user decide what
  should be the correct interpretation
  \item Performance: response time / evaluation metrics
\end{itemize}

\subsection{Data}
\label{sec:data}
%\TODO{Discuss the different types of structured data models that have been
% used, make clear that while they vary in expressiveness, they commonly enables
%capturing resources/entities/objects and relationships between them. Besides
%domain information captured in terms of entities and relationships, also point
%out other popular types of information that have considered, e.g. temporal and
%spatial information}

There is a great variety of structured data sources:
\begin{itemize}
  \item early data structures (Prolog databases; hierarchical lists)
  \item relational databases
  \item XML databases
  \item ontologies; \emph{linked data}
\end{itemize}
Structures of these data sources are different, as well as the way to access and
modify data (i.e. how to query data, which language to use, etc.).
For example, relational databases are represented with the entity/relationship
model and queries are usually expressed in a SQL-like language. Hierarchical
lists are hierarchical structures (like XML documents) that are composed of
embeded key-value pairs with optional attributes. XML documents are more
generic, since ``keys'' in those documents are nodes (i.e. can be trees
themselves). 
Ontologies are not more generic than XML databases from the structure point of
view, but allow to express easily semantic relationships.

Dome data structures are dedicated to some types of information. For instance,
temporal databases are similar to relational databases, but capture \emph{time}
in addition. Indeed, in some domains, the truth value of a fact or of
a resultset is determined by a \emph{timestamp}, while in classic database facts
are true in general.

At the end, all these data structures manipulate \emph{entities} (i.e. named
properties of the database) and \emph{relations} between these entities.
Any database selection query can be reduced to a set of constraints on these
entities and relations.

\subsection{Users' questions}
%\TODO{Discuss different types user questions}

\label{sec:big-picture-question}
Traditional database management systems provide an interface that allows users
to query the data in a formal language (e.g., SQL).

Question typing is a research task within the Q\&A field.
This community has classified questions in \emph{facto\"id questions} (i.e.
questions that can be answered by an \emph{objective} fact, typically questions
staring with wh-words except ``why'') and \emph{complex questions} which
answers are usually \emph{subjective} to the answerer, and for which there
might be several \emph{true} answers, possibly contradicting each other.
Complex questions can be ``why''-questions, ``how''-questions or
\emph{definition}-questions.
Facto\"id question is defined by Soricut \&
Brill~\cite{Soricut:2006:AQA:1127331.1127342} as questions ``for which a
complete answer can be given in 50 bytes or less, which is roughly a few
words''.
Several finer classification have been proposed in specific domains or
applications. 
For instance, in the case of temporal questions, Saquete et
al.~\cite{Saquete:2004:SCT:1218955.1219027} have proposed the classification
reproduced table~\ref{tab:saquete-classification}.
\begin{table}
\centering
\begin{tabular}{ll}\hline
\multicolumn{1}{c}{\textbf{Question type}} &
\multicolumn{1}{c}{\textbf{Example}}\\\hline\hline
Single event temporal questions &
``When did Jordan close \\
 without temporal expressions & the port of Aqaba to Kuwait?''\\\hline 
 Single event temporal questions & 
``Who won the 1988 New \\
 with temporal expressions & Hampshire republican primary?''\\\hline
 \multirow{2}{*}{Multiple events temporal questions} & ``What did G. Bush do
 after\\
  & the U.N. Security Council\\
 \multirow{2}{*}{with temporal expressions} & ordered a global embargo on\\
  & trade with Irak in August 90?''\\\hline
  \multirow{2}{*}{Multiple events temporal questions} & ``What happened to world
  oil\\
   & prices after the Iraqi\\
  without temporal expressions & annexation of Kuwait?''\\\hline

\end{tabular}
\caption{Question classification for temporal databases from~\cite{Saquete:2004:SCT:1218955.1219027}}
\label{tab:saquete-classification}
\end{table}

\subsubsection{Keyword query}
A {\it keyword query} consists in a set of words
traditionally used in the context of document search (i.e. searching relevant
documents given a textual query).
The classic idea is to represent documents in a vector model, where the vector
is composed of terms of documents. Search engines have made this paradigm
popular.
Hearst~\cite{Hearst:2011:NSU:2018396.2018414} reports that the number of words
in queries over Web search engines trends to increase (the experiment compared
queries performed over a month in 2009 and in 2010): queries from 5 to 8 words
increase up to 10\% while queries from 1 to 4 words decrease up to 2\%.
It seems indeed that users are more and more aware of new capabilities of
state-of-the-art search engines, and are less reluctant to express their
information need in a more ``natural'' way.

@@

\subsection{Internal representation}
\label{sec:big-picture-syntactic}

@@

\subsection{Database queries}
\label{sec:big-picture-semantic}
\TODO{Discuss query languages; discuss the different types of queries expressed
by these languages, e.g. entity queries, relational queries, additional
constructs such as group by etc.}

@@

\subsubsection{Expressivity of database query}
\label{sec:big-picture-expressivity}

%\subsection{Results}
%\label{sec:big-picture-results}

NL interfaces produce database queries.
Those queries must be then be executed by underlying DBMS.
The execution of queries can lead to different failing states:
\begin{itemize}
  \item the query execution fails (the generated database query is not valid)
  \item the query execution lead to an empty result set
\end{itemize}
In these cases, the system may inform the user and/or suggest/rephrase the
query. This is especially performed in systems belonging to the feedback
approach (see section~\ref{sec:feedback-driven}).

\subsection{Domain-independence}

\subsection{Portability}
